---
title: "GW Frameworks and approaches"
author: "Lex Comber & Paul Harris"
date: "June2016"
output: ioslides_presentation
---

## Geographically Weighted Approaches

- GW regressions (Brunsdon et al. 1996) are similar to ordinary regression
- **BUT** they use a moving window and compute ***local*** regression models at locations throughout the study area
- User accuracy
$$P(y = 1) = logit(b_0 + b_1x_1)$$
- GW User accuracy 
$$P(y_i = 1) = logit(b_{0(u_i, v_i)} + b_1x_{1(u_i, v_i)}    (eqn 2) $$
where 
$x_1$ and $y_1$ are classified and reference data
$(u_i, v_i)$ is a vector of two-dimensional coordinate

## Geographically Weighted Approaches

- GW regressions (Brunsdon et al. 1996) are similar to ordinary regression
- **BUT** they use a moving window and compute ***local*** regression models at locations throughout the study area
- Producer accuracy
$$P(x = 1) = logit(b_0 + b_1y_1)        (eqn 3)$$
- GW Producer  accuracy 
$$P(x_i = 1) = logit(b_{0(u_i, v_i)} + b_1y_{1(u_i, v_i)}    (eqn 3) $$
where 
$x_1$ and $y_1$ are classified and reference data
$(u_i, v_i)$ is a vector of two-dimensional coordinate

## Kernel, Moving window 

- GW approaches use data under the kernel to construct a (local) model at each location 
- Data are weighted with points further away contributing less 
- The weight, $w_i$ atlocation $(u_i, v_i)$ is a decreasing function of $d_i$, the distance to the kernel centre

- the Kernel need a shape
- eg the *bisquare* kernel
$$
f(d) = \left\{ \begin{array}{cl}
  \left(1 - \left( \frac{d}{h} \right)^2 \right)^2 & \mbox{ if $d < h$}; \\
   & \\
  0 & \mbox{ otherwise.} \end{array} \right.
$$

## Kernel, Moving window 
- The kernal size or **bandwidth** $h$ can be ***fixed*** or ***adaptive***
- Larger values of $h$ -> greater smoothing 
- In this work we define the bandwidth include the nearest 15% data points


## Load the data and packages
```{r eval=T,  message=F, warning=F}
library(GISTools)
library(spgwr)
library(repmis)
source_data("http://github.com/lexcomber/LexTrainingR/blob/master/SpatialAccuracy.RData?raw=True")
```

## Create spatial data  
- Indicates where the Field data / class is equal to the Classified data / class 

```{r eval=T,  message=F, warning=F}
res <- vector(length = dim(data)[1])
for (i in 1: dim(data)[1]) {
	if (data$Boolean_RS[i] == data$Boolean_FS[i]) {
		res[i] <- 1
	}}
res.spdf <- SpatialPointsDataFrame(coords = data[,2:3], 
  data = data.frame(res))
```

 - Define the a logit function:
```{r eval=T,  message=F, warning=F}
alogit <- function(x){exp(x)/(1+exp(x))}
```

## Create a grid over which to pass the kernel 
```{r eval=F}
grid <- SpatialPoints(expand.grid(x=seq(295000,363000,by=1000),
  y=seq(3610000,3646000,by=1000)))
par(mar = c(0,0,0,0))
plot(roilib)
plot(grid, cex = 0.7, col = "grey")
plot(res.spdf, add = T, pch = 1, col = "#25252580")
```

## Create a grid over which to pass the kernel 
```{r echo = F, eval=T, fig.height = 5}
grid <- SpatialPoints(expand.grid(x=seq(295000,363000,by=1000),
  y=seq(3610000,3646000,by=1000)))
par(mar = c(0,0,0,0))
plot(roilib)
plot(grid, cex = 0.7, col = "grey", add = T)
plot(res.spdf, add = T, pch = 1, col = "#25252580")
```

## GW overall accuracy example
```{r eval=T,  message=F, warning=F}
bw = 0.15
gwr.mod <- ggwr(res~1, data = res.spdf, adapt = bw, 
  fit.points = grid, family= binomial) 
gwr.ov <- alogit(data.frame(gwr.mod$SDF)[,2])
```
- The results show only a small amount variation: 
```{r eval=T, message = F}
summary(gwr.ov)
```

## GW overall accuracy example
```{r, echo = F, eval=T, message = F, fig.height = 6}
shades = auto.shading(gwr.ov, cols=brewer.pal(4,"Greens"))
gwr.res = SpatialPixelsDataFrame(gwr.mod$SDF, data.frame(gwr.ov))
par(mar = c(0,0,2,0))
level.plot(gwr.res,shades)
lib.masker = poly.outer(gwr.res, roilib, extend = 100)
add.masking(lib.masker) 
plot(roilib, add = T)
choro.legend(297000, 3650000,shades) 
```

## Summary
- Geographically Weighted approaches are used toconstructs *local* models of accuracy
- References
    - Brunsdon, C.F., Fotheringham, A.S. and Charlton M. (1996). Geographically Weighted Regression - A Method for Exploring Spatial Non-Stationarity, *Geographic Analysis*, 28: 281-298.
    - Gollini, I., Lu, B., Charlton, M., Brunsdon, C., and Harris, P. (2013). GWmodel: an R Package for Exploring Spatial Heterogeneity using Geographically Weighted Models. *arXiv preprint arXiv*:1306.0413.

