---
title: "Introduction: the spatial analysis of accuracy using geographically weighted frameworks"
author: "Lex Comber & Paul Harris"
date: "June 2016"
output: ioslides_presentation
---

## Introduction

- In this first session you will  

    - Open a point dataset (shapefile) of a field validation exercise
    - Inspect the attributes and spatial properties of the data 
    - Generate a correspondence matric and some classic measures of accuracy
    - Generate some surfaces

## Introduction

- A zip file of the worksheet `PDF` and `Rmd` files for this workshop can be downloaded:  
- [https://github.com/lexcomber/AccuracyWorkshop2016/blob/master/GW_Accuracy_Data.zip](https://github.com/lexcomber/AccuracyWorkshop2016/blob/master/GW_Accuracy_Data.zip). 
- Save this to your working directory 
- It can be loaded directly.
- You will need to load some R packages: 
```{r eval=F}
install.packages("repmis", dep = T)
install.packages("GISTools", dep = T)
install.packages("spgwr", dep = T)
```

## Introduction

 - ...and then call the packages (or *libraries* 

```{r eval=T,  message=F, warning=F}
library(GISTools)
library(spgwr)
library(repmis)
```
 - and load some data   
```{r eval=T,  message=F, warning=F}
source_data("http://github.com/lexcomber/LexTrainingR/blob/master/SpatialAccuracy.RData?raw=True")
```

## Introduction

- And have a look at the what you have
```{r eval=T}
ls()
data.frame(data)[1:2,]
```

## Introduction

- The number of continuous and categorical variables
    - First few columns: ID (`PointID`) X and Y 
    - Next set of 5 variables (columns 4 to 8), all with the suffix `_FS` are the fuzzy memberships 
    - We will not use use these!
- We will work with 
    - `Boolean_FS`- *Field Survey* class, the *Observed* or *Reference* data
    - `Boolean_RS`- *Remote Sensing* class the *Predicted* or *Classified* data
- We will mainly be using the Boolean data in columns 9 and 15

## Map the data
 - We will map the data 
    - Creating a `SpaatialPointsDataFrame` class of object
    - Defined in the `sp` package 
    - Loaded with the `GISTools` package

```{r eval=T}
# Convert data to SPDF - spatial data
# 1.define a projection - see http://spatialreference.org/ 
lyb.proj <- 
  CRS("+proj=utm +zone=33 +ellps=WGS84 +datum=WGS84 +units=m +no_defs ") 
# 2. then use this to project the data from its coordinates
data.spdf <- SpatialPointsDataFrame(coords = data[,2:3], 
  data = data, proj4string = lyb.proj)
### define Tripoli
x <- 329266.6; y <- 3640653; loc <-cbind(x,y)
Tripoli <- SpatialPointsDataFrame(coords = loc, data = data[1,])
```

## Map the data
```{r eval=T}
### Map
par(mar = c(0,0,0,0)); plot(roilib)
plot(Tripoli, pch = 19, add = T, cex = 3, col = "#31A354")
plot(data.spdf, add = T, pch = 20, cex = 1, col = "Black")
```

## The correspondence matrix
 - We will be generating local models of the ccorrespondenc matrix 
    - AKA the *accuracy*, *confusion*, *error* or *validation* matrix
 - Overall accuracy: from the diagonal and off-diagonal elements
 - User and producer accuracies from the diagonals and  marginal row and column totals 
- Create the correspondence matrix: 
```{r eval=T}
tab <- table(data$Boolean_RS, data$Boolean_FS)
```
 - This generates a table of *Predicted* or *Classified* (rows) against *Observed* (columns). 

## The correspondence matrix
```{r eval=T}
class.names.long <- c("Bare", 
    "Grazing", "Urban", "Vegetation", "Woodland") 
rownames(tab) <- class.names.long
colnames(tab) <- class.names.long
tab <- cbind(tab, rowSums(tab))
tab <- rbind(tab, colSums(tab))
rownames(tab)[6] <- "Total"
colnames(tab)[6] <- "Total"
tab
```

## Accuracy measures

- Overall User and Producer accuracies 
- See Congalton (1991)
    - [http://uwf.edu/zhu/evr6930/2.pdf](http://uwf.edu/zhu/evr6930/2.pdf). 
```{r eval=T, echo = F}
# Users accuracy
tmp <- vector(mode = "numeric", length = 6) 
for (i in 1:5) {
	tmp[i] <- tab[i,i] / tab[i,6]
	}
tab <- cbind(tab, zapsmall(tmp, 3))
colnames(tab)[7] <- "Users"
# Producers accuracy
tmp <- vector(mode = "numeric", length = 7) 
for (i in 1:5) {
	tmp[i] <- tab[i,i] / tab[6,i]
	}
tab <- rbind(tab, zapsmall(tmp, 3))
rownames(tab)[7] <- "Producers"
tab[7,7] <- sum(diag(table(data$Boolean_FS, 
  data$Boolean_RS)))/sum(table(data$Boolean_FS, data$Boolean_RS))
```

```{r eval=T}
round(tab, 2)
```

## Correspondence, Probabilities and Regression
-  *Overall*, *User* and *Producer* accuracies are probabilities
- These can be calulated using logistic regressions
- By predicting the degree to which the data (the accuracy meausure) is equal to 1  
- Detail in the worksheet
- In the next part you will map these 

