---
title: "Mapping Spatial Accuracy"
author: "Lex Comber & Paul Harris"
date: "June2016"
output: ioslides_presentation
---

## Introduction

- You have created a spatially distributed measure of **overall** accuracy
    - Whether the *observed* class = *predicted* class
- Using logistic regression (Generalized Linear Models) 
- An *a*spatial measure then GW logistic Regression 
- Now you will develop additional measures 
    - User and Producer accuracies 
- A full description on Congalton (1991):   
    - [http://uwf.edu/zhu/evr6930/2.pdf](http://uwf.edu/zhu/evr6930/2.pdf).
- You will calculate these first principles and then develop  *GW* versions

## Data set up

- As before, you will need to load the data. 

```{r, echo = F, eval=T,  message=F, warning=F}
library(GISTools)
library(spgwr)
library(repmis)
source_data("http://github.com/lexcomber/LexTrainingR/blob/master/SpatialAccuracy.RData?raw=True")
```

- And use the code from the *Introduction* to create the correspondence matrix

```{r, echo = F, eval=T}
tab <- table(data$Boolean_RS, data$Boolean_FS)
class.names.long <- c("Bare", "Grazing", "Urban", "Vegetation", "Woodland") 
rownames(tab) <- class.names.long
colnames(tab) <- class.names.long
tab <- cbind(tab, rowSums(tab))
tab <- rbind(tab, colSums(tab))
rownames(tab)[6] <- "Total"
colnames(tab)[6] <- "Total"
```
 - And calculate Overall, User and Producer accuracies 
```{r, echo=F,eval=T}
# Users accuracy
tmp <- vector(mode = "numeric", length = 6) 
for (i in 1:5) {
	tmp[i] <- tab[i,i] / tab[i,6]
	}
tab <- cbind(tab, zapsmall(tmp, 3))
colnames(tab)[7] <- "Users"
# Producers accuracy
tmp <- vector(mode = "numeric", length = 7) 
for (i in 1:5) {
	tmp[i] <- tab[i,i] / tab[6,i]
	}
tab <- rbind(tab, zapsmall(tmp, 3))
rownames(tab)[7] <- "Producers"
tab[7,7] <- sum(diag(table(data$Boolean_FS, 
  data$Boolean_RS)))/sum(table(data$Boolean_FS, data$Boolean_RS))
```

## Data set up
```{r eval=T}
round(tab, 2)
```

## Recap: User Accuracy
- Provides a measure of per-class accuracy 
- Describes how probable it is that a pixel (or segmented object) of that has been labelled as *Grass* will actually be that class if it was visited in the field 
- Indicates the errors of commission (inclusion) 
- For the potential **user** of the map 
    - Probability of correctly finding the class indicated on the map present on the ground

## Recap: User Accuracy

- Recall that the approach for User accuracy was as follows:
    - For class `x`, create a `data.frame` containing: 
        - Locations where the remote sensing indicated class `x` [0 or 1]
        - Locations where the field visit indicated class `x` [0 or 1]
- Construct a GLM of the extent to which the field class was predicted by the remote sensing class (*Observed* was predicted by *Predicted*)
- Manipulate the resulting coefficients (`sum` and the `alogit` function) to determine the User accuracy (remembering that User accuracy is $P(FS = 1|RS = 1)$

```{r echo=F, eval=T}
# 1. Create a data.frame
class.list <- unique(data$Boolean_RS)[order(unique(data$Boolean_RS))]
# 'G' is for Grazing Land  
class <- class.list[2]	
# 1a where the RS indicated the class
rs.class <- (data$Boolean_RS == class) * 1
# 1b where the FS indicated the class
fs.class <- (data$Boolean_FS == class) * 1
# join together
fsrs <- data.frame(cbind(fs.class,rs.class)) 
```

## GW User Accuracy
- Construct a `SpatialPointsDataFrame` object
- Input to the GW GLM function, `ggwr`:
```{r eval=F}
fsrs.spdf <- SpatialPointsDataFrame(coords = data[,2:3], 
  data = data.frame(fsrs))
grid <- SpatialPoints(expand.grid(x=seq(295000,363000,by=1000),
  y=seq(3610000,3646000,by=1000)))
plot(grid, cex = 0.7, col = "grey")
plot(roilib, add = T)
plot(fsrs.spdf, add = T, pch = 1, col = "#25252580")
```

## GW User Accuracy
- Construct a `SpatialPointsDataFrame` object
- Input to the GW GLM function, `ggwr`:
```{r echo = F, eval=T}
fsrs.spdf <- SpatialPointsDataFrame(coords = data[,2:3], 
  data = data.frame(fsrs))
grid <- SpatialPoints(expand.grid(x=seq(295000,363000,by=1000),
  y=seq(3610000,3646000,by=1000)))
plot(grid, cex = 0.7, col = "grey")
plot(roilib, add = T)
plot(fsrs.spdf, add = T, pch = 1, col = "#25252580")
```

## GW User Accuracy
- Specifiy a bandwidth of 15% 
```{r eval=T}
bw = 0.15	
```
- And construct the GW model can be constructed.
```{r eval=T,  message=F, warning=F}
gwr.mod <- ggwr(fs.class~rs.class, data = fsrs.spdf, 
  adapt = bw,fit.points=grid, family= binomial) 
```

## GW User Accuracy
- And examine the (`gwr.mod`) 
```{r eval=T}
gwr.mod
```

## GW User Accuracy
- The GW coefficients can be manipulated 
```{r eval=T}
coefs <- data.frame(gwr.mod$SDF)[,2:3]
coefs[,2] <- rowSums(coefs) 
# P(x = 1|y = 1)
alogit <- function(x){exp(x)/(1+exp(x))}
gwr.user <- alogit(coefs[,2]) 
```
- Spatial variation - compare the 1st and 3rd quartiles
```{r eval=T}
summary(gwr.user)
```


## GW User Accuracy
- This can be mapped
```{r echo=F, eval=T, fig.height = 6}
shades = auto.shading(gwr.user, n=5,cols=brewer.pal(5,"Greens"),
  cutter=rangeCuts, digits = 2)
gwr.spdf = SpatialPixelsDataFrame(gwr.mod$SDF, data.frame(gwr.user))
par(mar = c(0,0,1,0))
level.plot(gwr.spdf,shades)
lib.masker = poly.outer(gwr.spdf, roilib, extend = 100)
add.masking(lib.masker) 
plot(roilib, add = T)
choro.legend(297000, 3650000,shades) 
title("User Accuracy: Grazing Land")
```

## GW User Accuracy

```{r echo=F, eval=T}
# 2. GLM for User Accuracy
mod1 <- glm(fs.class~rs.class,data = fsrs,family= binomial) 
mod.coefs <- mod1$coefficients
mod.coefs[2] <-sum(mod.coefs) 
# P(y = 1|x = 1)
mod.user <- alogit(mod.coefs[2])
```
- The *global* measure of User accuracy was `r round(as.vector(mod.user), 3)` and *locally* this varies from `r round(summary(gwr.user)[2],3)` to `r round(summary(gwr.user)[5],3)`:

```{r eval=T}
# global
round(as.vector(mod.user), 3)
# local
round(summary(gwr.user)[c(2,5)],3)
```
- Producer Accuracy is dealt with in a similar way 


